Integrating Social Visual Analysis and Deployment with RCloud

### 2015-03-28 interviews.tex
further narrate the interviews

\gordon{Add user statistics here: number of notebooks, forks, stars, users
who have done one of these things. Number of users who use RCloud a lot,
a moderate amount, or just tried it out.}

@@

Besides providing a way to present work to others, notebook sharing can also
provide a starting point. Clint says, ``I think the best part about it is how
easily you can share code... you can find a working example, rather than wearing
out Google and finding questionable examples that may or may not work.'' Raif
notes, if ``some person has done something similar, then you're able to just
edit that, and that's saved a lot of work time for me''

Taylor develops packages for analysts: ``I use it more for sharing code with
other people, and for doing tutorials for iotools or hmr'' or explaining issues
with data sources. ``A lot of times I'll share the code because I'm doing so
much package development. I want people to see how the package works, so I
clearly want them to see the code... I kind of write it just like I would write
GitHub Markdown, where you have little code snippets and text, but RCloud lets
me actually run the snippets'' and display the results.

Some users who tried out RCloud but are not able to use it for organizational
reasons, already miss the functionality. Shelly likes
``the concept of being able to create notebooks and share them.'' A ``wiki is not the
best way for communication results - it's kind of like writing a blog post with
very limited functionality... I have to save every
picture and post it as an image'' and ``I can't share all of the code because it
would just get crowded and wouldn't look right on a wiki.'' Bart says, ``if I could
make a folder on
RCloud and have Python notebooks and also Pig notebooks there, and execute them
from RCloud, that would be much better than my current thing, because that would
free me from manual documentation and version control and also telling people
where my code was. It would be just, hey, go look on RCloud, here's the stuff.''

@@

\subsection{Forking}
Forking is another popular feature. 
Clint: ``I use [forking] on my own work quite a bit, and I use it on Research a
ton. Honestly, it's one of the handiest functions, because instead of
having to find it, copy, paste it, you just hit Fork, you rename it, and it's
done. It's pretty amazing.''

@@

A somewhat more problematic and extremely common use of forking is just to change parameters. Raif: ``I've been forking other people's notebooks myself, because essentially
they have done some work, but then I want to run it on a different part of data,
or I want to change some parts, I don't want to see this column, I want a
different column, things like that.'' Horace: on working with other group using his notebooks, ``I basically point them to a few different parameters that they need to fix.''

But forking also provides a way for others to troubleshoot when something goes wrong.
Horace continues, ``I'll teach people
to intercept the result in the middle and see what you should see, and what's
not being returned, that sort of thing... If the computation went wrong in the
middle, and something blew up, or you're just returning NULL or empty data,
where should we check those things?''

@@

\subsection{Automatic source control}
Rick: ``when I've finished something, there's a nice clean record of it.''

Clint: ``instead of looking back and saying I've got a billion files here in
this subdirectory and I hope I've got them backed up, if they're on RCloud I
know they are'' ``I don't have to worry about it if I drop my computer in the
bathtub, everythings gone, because I'm not great about uploading my own for
versioning''

Bart: ``I like the fact that it has a built-in editor, so if you need to
fix a typo in a link, or an extra line break or any of this other nonsense,
you can just switch to the edit view, pull up your asset, type something, it's
automatically saved, committed, everything. You don't have to go back to your
source code, change it, commit it to the repo, pull the repo to your
distribution version, none of that's necessary. A lot of web development is
fixing stupid little bugs, and you can do that instantly without a lot
of overhead, which is nice.''

\subsection{Search versus Browsing}
Raif: ``the fact that we have all the notebooks there, searchable, it saves
me from replicating what other people have done'' ``One way of thinking of it is
like Google Code, you know you just search for something and you find something
useful and you just use it. It’s a repository of code''

But global search doesn't work for everyone, and may need to be refined. Other users
prefer to just browse the notebooks of the experts they know. Rick says, ``usually I
know somebody’s notebooks that I want to search through, like I’ll search
through Simon Urbanek’s first, then I’ll go off to somebody else’s, because I
know that the kind of thing I’m looking for something more obscure than I’m
likely to find in some random person’s thing''

\subsection{Web integration}
Horace ``that also means the other guys who want to do analytics on the data,
they can first pull the data, do the analytics on it, and then feed the viewer
the data. So if anyone from the stats group wants to insert some code in between
the data pulling procedure, [they] can always insert something in between. We
were trying to make such an example so that they don't have to rewrite the GUI
part and they know how to get the data source, once you get the data into
RCloud, then you have a dataframe to work with, and then they know how to
produce another dataframe. We give them the specification of the dataframes, we
give them the data input, they do whatever they want, they can reuse the data
connection part and the GUI part. We can keep changing these components and
that's a very nice thing.''

Jyothi: ``Having an open session with R certainly helps too, because both of my
applications need to be doing something with the data, so having an open session
where I can run R command or R functions without having to invoke an API or send
out a request and then wait for the request to come back to do the results.  So
that kind of seamless integration is extremely helpful in writing the
application.''

Jyothi: ``another thing I really like about RCloud is your publish feature, so
that non-RCloud users, the anonymous login, non-RCloud users can access the
notebooks without having to login to RCloud''

accidental feature: ``some of my jobs are huge and I like to just call them and
the fact that the user can just shut down his browser or shut down his computer
and the job still runs in the background is actually a very neat thing. So what,
um, the way I have coded is, whenever I have a huge job to run, I just display
to the user, you go do whatever you want, we'll send you an email when the job
is done, um that's another feature of RCloud that's very helpful, that the job
will continue to run in the background''

\subsection{Other workflows}
Rick: ``I find that the style of editing is clunky. I would never edit like that
if I'm doing ordinary work. I usually use vi.'' ``I mean it's nice to have it
saved, but you know, there's sort of this trade-off between it making it easier
for me to present something or to save something, and my easy of typing and
correcting and things in a plain old editor window.'' ``I just have the
discipline of having a file that describes what I'm doing with the commands that
I'm using so that I can go back and recreate it or pass it on to someone else,
so it's a little bit like having an RCloud notebook, it's not necessarily
executable but most of the commands that I've typed are in there.''

Kenny: the UI doesn't work for his workflow, which is pasting snippets from a
file into the console in the R GUI.  he tends to type type type into the file,
paste paste paste to the console, rearrange so the good code is at the top of
the file, then save the file and call it quits once he's got the right code at
the top of the file.

Taylor: ``The big one is I don't know why would I want to use RCloud over my
current setup. If it's just me, I like my text editor and terminal. There's
nothing that I want that those two don't give me'' ``In R, when I'm doing
development work, I copy and paste in like 5-10 lines of code, so when something
breaks, I get an error message on that one line, and I can up-arrow and change
it and fix it, whereas in RCloud I have to run a whole cell, so the only way to
get that same functionality is if every line's in one cell. Because if it runs
like lines 1,2,3, then 4 fails, I have to either run the whole cell again, or
have line 4 in its own cell. I can't rerun just one line.'' ``There's nothing
that I want that's not in the terminal. I don't think you should try to support
all the things I want to do in the terminal.''

Horace: ``It's not for myself, but for other people who use RCloud vs RStudio,
they will always think that they want to prototype on RStudio. Maybe because
it's more familiar for them. So, they work on RStudio, on a small data set, and
they fix up everything, make sure everything is right, and then they place it on
RCloud for sharing.''

Byers: installing random packages is part of exploratory data analysis

\subsection{Proliferation of notebook cells}
Rick: ``Making new cells and taking some stuff in it, then having it do
something, then having to make a new cell, and I don't know, it just gets in my
way, so if I'm doing really straightforward simple stuff, I'll just get into R
and do it myself.'' ``I'm always scrolling up and down the screen'' ``every time
I type more stuff, the notebook gets longer and longer and it's harder to deal
with'' ``I've got to do a lot of scrolling around to get to where I'm going,
whereas if I'm doing it in vi, I can search for things in my typing, and get to
places I want to be without having to scroll through it and try to read it''
``times when I just want to type in a couple of quick and get some results that
are going to tell me what to do next, and they're not necessarily archival in
any sense'' ``being literate about anything typically takes a lot of rewriting
and going back over things... I just want to explore a few things and then I'll
know what I want to write.'' ``I'll often type in expressions that allow me to
check that I'm the right track, so I know that if the computation worked out,
then the sum of the residuals is gonna be zero. So I'll type in sum(resid) and
if it prints out zero, then I'm all happy, but I don't want that in my
notebook.'' ``it tempts me to type a big long thing and then run the whole
thing, as opposed to, type a few little pieces that I'm thinking of on the inner
parts and then put them together to make the big thing''

Ivan has the complaint that too many cells are created to use the notebook as a
console.  This has a lot to do with the extra real estate that gets taken up by
the cells controls, blank space, etc of cells versus a straight command line.

KC and Ivan both complained that when there are long results or plots, it causes
the code to scroll off the screen.

Kenny: ``It saves everything I do like everything is gold, but most of it is
junk not meant to be saved'' ``the console is purely disposable''

Raif: ``whatever you type on the command line, it becomes a part of your
notebook, and I find it a little bit annoying, because I really type a lot of
things to check my data, and I don't want that to become a part of the
notebook'' ``Like you're kind of checking out what your data is, or you make a
plot of the data. Things that should not really become part of a notebook, but
things that help you understand your data better.''

Horace: ``Well, the nice thing is if you have the whole notebook, you can run it
step by step, and try to mess around in between, interactively.''

@@

\subsection{Misc. notebook interface annoyance}
Rick: ``if I was typing into one of the cells near that top, I had to think of
`I'm editing this cell and then I have to execute it', and if was doing the one
at the bottom I could type it but then it would automatically execute, but then
it became a standard cell and I have to edit it. It was just another set of
modes that I didn't like''

Clint: ``it took me a couple of weeks of looking at it to become comfortable
because I saw the cells and it just threw me for a total loop. I mean, it's a
good idea because it's essentially like I can have this part and this part and
this part. I can run it all at once if I want or I can break these into sections
for either debugging or staging purposes which, I really like it now, but when I
first saw it, [it was] very confusing... I think I prefer it now, because you
can cut your code up.''

Corey: In the previous version, it switched between code and results. Now you
see both - ``and there are times when that's good, but there's certainly times
when it'd be nice to - you know, you have 10 or 15 cells and you just want to
get back and there's all this output and you're scrolling all over the place so,
if there were a way to control both, that'd be great. There are times when I do
want to see the output, but there are times when I've seen enough of the same
thing.''

Jyothi: ``It's good that I can see the results but if I'm working and I need
more room and I want to hide the results, I can't.'' Could we have ``an option
to close the results window?''

Ivan: RStudio's layout is more helpful because - the charts are kept in another
pane and stay in one place while doing analysis - having ``long term'' code
(which would be in assets) above the command prompt means that both can be sized
very wide to accomodate long lines

Raif: ``cells are really useful'' but ``you want to see your output in a
different window, on a separate part of the window'' ``for example imagine I run
a cell whose output is huge, right, and now that output would block the whole
window, and in order to find the next cell, I have to scroll down, and find
where that cell starts, right? So, I'm losing the continuity of my code to the
output''

Kenny: Problem with screen real estate: side panes take up too much
space. Doesn't need the notebook tree pane.

KC thought that having a scratch pad that doesn't persist between different
notebooks is ``useless''.

Kenny: Problem of long-running cells.  ``The long tail.''  First cell takes
seconds to run, next cell minutes, next cell 5 hours.  In this scenario, which
is common for Kenny, the ``Run whole notebook'' button is dangerous. This is
also a huge issue for demoing code.  Suggests something like knitr's caching of
results. When converting his work to notebooks, Kenny ends up with a lot of
comments that say ``this cell takes a long time to run.'' He ends up ``littering
the notebook with little switches that comment out'' the slow parts and ``load
the object instead'', ``save the object to disk''.

\subsection{A sea of notebooks}

Byers: Problem with forking a notebook and correcting it, is that the original
notebook still exists with the error.  We need ways to see that there is a more
recent fork of notebook and to fetch in the changes from it.

Taylor: ``I'll make a notebook of a little tutorial of, here's the basic
functions and what they do, and then I'll send them out, and then people will
give me feedback on them, or have some new feature, or I'll realize something's
wrong, and then I'll go back and change it and update that page, but then what
happens is people have in the meantime, forked it, and then they have a version
that... and usually what it'll be is actually that I'll have to change a package
as well, so their old fork stops working, and then they complain, and then, oh
no, you have to go get the original [notebook]. Yeah, that's the sort of... This
probably bites me more than other people, because I'm doing development so
things are changing as they're looking at the code. Like, iotools, a lot of the
functions change names or change syntax.''

Taylor: ``That's just a problem because people will have old notebooks that
don't work anymore. It's annoying when it's a notebook that I still have and the
format has changed just a bit. More than that is when people have, now that I
know that you can't even delete notebooks, they'll have a fork of something that
doesn't exist anymore on my thing.'' ``Now I'm really gun-shy about sharing
notebooks because, I don't want to like, it's like, do I want to support this
forever?''

Taylor: ``A good example is, someone was complaining about doing forking in an
RCloud notebook. They were trying to use a parallel package. And I said, no no I
think it works, and I did like a one-cell thing that just did some sweep kind of
thing to show that it was actually working in parallel, and then I just deleted
it, thinking it was gone. And then it saved something in a directory that I then
deleted, because it was just a temporary directory. Three months later, I don't
know why they were running it, they said `oh it doesn't work'. I think that's
the biggest issue I'm having right now, all these forks''

Taylor: ``there's no way for both of us to have ownership of a notebook, so the
only way is to fork it back and forth, and so we have dozens of old copies. [We
  end up] deleting all the old ones, but people still have links to them,
because they don't actually disappear.'' ``The UI makes it easy to go back to
the old stale versions, because you can say `forked from' `forked from' `forked
from', but there's actually no way to go the way that we want people to go,
which is, like, up the tree.''

Rick: ``I don't necessarily need to see everybody's notebook that uses RCloud''
``every time I do something new, I get a new notebook and so now my notebooks
are maybe 50 or 60. That's enough to think about just on my own, but if I've got
everybody's 50 or 60 sitting on my display, I find that it's more than I want to
know about.''

\subsection{Versioning not useful}
Rick: ``I don't need something that keeps track of every mistake I've made or
every direction I've tried.'' ``namespaces are already crowded, so trying to
remember the names of notebooks is hard, much less the names of states within
each notebook.''

\subsection{Web snafus}
Corey: ``What frustrates me is that if I do leave this window for what feels
like 5 or 7 minutes, it disconnects and I have to reconnect and run the whole
notebook again.''

Kenny: it really sucks when you get disconnected

Raif: ``being web-based, it's obviously slower than something you would be
running on your machine''

Byers: even the little lag of sending a command to a server and receiving the
result is intolerable

\subsection{Parameterization}
Clint: ``For the things that I'll do multiple times, I'll try to parameterize
it, and I'm working now on putting some packages together that will operate in
some fashion outside of RCloud, straight from the command line''

Taylor: ``like the top of it would be, `This is a report of the volume of all of
our feeds for this month', and someone would want to look at it for the next
month or the previous month, so they'd fork it to change the month''

Jyothi: ``I think that if a notebook is web-enabled [parameterized] then it
gives the person who is running the notebook, I mean they can always fork the
notebooks and make changes, but I feel that if the owner of the notebook
web-enables it, then whoever is running it, it's easier to run, instead of
forking it, if they can set options, it's probably more efficient, and also they
can fork it if they want to.''

\subsection{Other observations}
Byers: There is also a big problem with ``bitrot'' - notebooks often stop
working because the user changed the structure of their data, changed a filename
or a database.  We need ``organizational protocols'' to catch up with the
technology.  There is a ``hysteresis between the technical and the social'' and
we have not yet adapted.

Taylor: `` It would be a file inside a directory that only I had read access to
and so someone else couldn't run it... That was usually what it was, it was a
temporary file on something that was read-only to everyone else. But I've gotten
careful about that.''

Bart: ``One UI notebook and a bunch of other notebooks that do various tasks and
are called with API calls. The problem is, if you clone all the notebooks, then
you have to go update every single notebook, because all the IDs are different
now. There's no sort of relative path type stuff that you can do. You can do it
with static assets, but you can't do it with Ajax calls''

>>>>

@@

Besides providing a way to present work to others, notebook sharing can also
provide a starting point. Clint says, ``I think the best part about it is how
easily you can share code [...] you can find a working example, rather than wearing
out Google and finding questionable examples that may or may not work.'' Raif
notes, if ``some person has done something similar, then you're able to just
edit that, and that's saved a lot of work time for me''

Taylor develops packages for analysts: ``I use it more for sharing code with
other people, and for doing tutorials for iotools or hmr'' or explaining issues
with data sources. ``A lot of times I'll share the code because I'm doing so
much package development. I want people to see how the package works, so I
clearly want them to see the code [...] I kind of write it just like I would write
GitHub Markdown, where you have little code snippets and text, but RCloud lets
me actually run the snippets'' and display the results.

Some users who tried out RCloud but are not able to use it for organizational
reasons, already miss the functionality. Shelly likes
``the concept of being able to create notebooks and share them.'' A ``wiki is not the
best way for communication results - it's kind of like writing a blog post with
very limited functionality [...] I have to save every
picture and post it as an image'' and ``I can't share all of the code because it
would just get crowded and wouldn't look right on a wiki.'' Bart says, ``if I could
make a folder on
RCloud and have Python notebooks and also Pig notebooks there, and execute them
from RCloud, that would be much better than my current thing, because that would
free me from manual documentation and version control and also telling people
where my code was. It would be just, hey, go look on RCloud, here's the stuff.''

@@

\subsection{Forking}
The ability to start work where someone else let off, by forking, has proved to
be a popular feature. In fact, almost twice as many (131) users have forked
someone else's notebook as have starred (75).  Clint: ``I use [forking] on my
own work quite a bit, and I use it on Research a ton. Honestly, it's one of the
handiest functions, because instead of having to find it, copy, paste it, you
just hit Fork, you rename it, and it's done. It's pretty amazing.''

@@

Forking also provides a way for others to troubleshoot when something goes
wrong.  When Horace works with the users of his notebooks, ``I'll teach people to intercept the result in the
middle,'' to ``insert print statements here and there and check values''

@@

An extremely common but more problematic use of forking is to
change parameters. Raif: ``I've been forking other people's notebooks [because]
I want to run it on a different part of data, or I want to change some parts, I
don't want to see this column, I want a different column, things like that.''

Our users, especially those who build general purpose tools, see a need for better ways
to parameterize notebooks.  (Currently parameters can be added to the URL, but adding
user interface elements to do the same thing requires expertise.) Taylor complains that
a notebook might say `` `This is a report of the volume of all of
our feeds for this month', and someone would want to look at it for the next
month or the previous month, so they'd fork it to change the month.''

Jyothi calls it web-enabling the notebook: users ``can always fork the
notebooks and make changes, but I feel that if the owner of the notebook
web-enables it, then whoever is running it, it's easier to run. Instead of
forking it, if they can set options. It's probably more efficient, and also they
can [still] fork it if they want to.''

%% Clint: ``For the things that I'll do multiple times, I'll try to parameterize
%% it, and I'm working now on putting some packages together that will operate in
%% some fashion outside of RCloud, straight from the command line''

\subsection{Automatic source control}
%% Rick: ``when I've finished something, there's a nice clean record of it.''
%% (I think this may be more about the publish/markdown than about source control.)

Clint: ``instead of looking back and saying I've got a billion files here in
this subdirectory and I hope I've got them backed up, if they're on RCloud I
know they are [...] I don't have to worry about it if I drop my computer in the
bathtub, everythings gone, because I'm not great about uploading my own for
versioning''

Bart points out that the automatic versioning works well for dealing with the
minutiae of web development. ``I like the fact that it has a built-in editor, so
if you need to fix a typo in a link, or an extra line break or any of this other
nonsense, you can just switch to the edit view, pull up your asset, type
something, it's automatically saved, committed, everything. You don't have to go
back to your source code, change it, commit it to the repo, pull the repo to
your distribution version. None of that's necessary. A lot of web development is
fixing stupid little bugs, and you can do that instantly without a lot of
overhead, which is nice.''

On the other hand, saving every change leads to many fine-grained
versions. Rick says that for this reason, the history feature is not all that
helpful: ``I don't need something that keeps track of every mistake I've made or
every direction I've tried.'' Tagging these versions with names wouldn't
necessarily help because ``namespaces are already crowded, so trying to remember
the names of notebooks is hard, much less the names of states within each
notebook.''

\subsection{Discovering others' work}
Many users find the search function useful. Raif says ``the fact that we have
all the notebooks there, searchable, saves me from replicating what other people
have done. [...] One way of thinking of it is like Google Code: you just search
for something and you find something useful and you just use it.''

But other users prefer to browse just the notebooks of the experts they
know. Rick says, ``usually I know somebody's notebooks that I want to search
through, like I'll search through Simon Urbanek's first, then I'll go off to
somebody else's, because I know that the kind of thing I'm looking for something
more obscure than I'm likely to find in some random person's.''

More selective ways to search will be needed as the corpus of notebooks
continues to grow. We are considering tagging notebooks filtering the notebook
tree by tag, measuring the quality of notebooks by their forks as well as stars,
and searching within a single user's notebooks.

\subsection{Web integration}
Integrating an analysis language into a web development environment is something
the tool developers really appreciate.  Horace: ``that also means the other guys
who want to do analytics on the data, they can first pull the data, do the
analytics on it, and then feed the viewer the data. So if anyone from the stats
group wants to insert some code in between the data pulling procedure, [they]
can always insert something in between. We were trying to make such an example
so that they don't have to rewrite the GUI part and they know how to get the
data source, once you get the data into RCloud, then you have a dataframe to
work with, and then they know how to produce another dataframe. We give them the
specification of the dataframes, we give them the data input, they do whatever
they want, they can reuse the data connection part and the GUI part. We can keep
changing these components and that's a very nice thing.''

Jyothi: ``Having an open session with R certainly helps too, because both of my
applications need to be doing something with the data, so having an open session
where I can run R commands or R functions without having to invoke an API or send
out a request and then wait for the request to come back to do the results,
that kind of seamless integration is extremely helpful in writing the
application.''

%% Jyothi: ``another thing I really like about RCloud is your publish feature, so
%% that non-RCloud users, the anonymous login, non-RCloud users can access the
%% notebooks without having to login to RCloud''

%% accidental feature: ``some of my jobs are huge and I like to just call them and
%% the fact that the user can just shut down his browser or shut down his computer
%% and the job still runs in the background is actually a very neat thing. So what,
%% um, the way I have coded is, whenever I have a huge job to run, I just display
%% to the user, you go do whatever you want, we'll send you an email when the job
%% is done. That's another feature of RCloud that's very helpful, that the job
%% will continue to run in the background''

\gordon{Not sure this belongs here}
Running anything on the Web can also be a source of frustration.  Poor network
links can cause web sockets to die, and Corey complains ``What frustrates me is
that if I do leave this window for what feels like 5 or 7 minutes, it
disconnects and I have to reconnect and run the whole notebook again.'' Raif
whines that ``being web-based, it's obviously slower than something you would be
running on your machine.'' For Byers: even the little lag of sending a command
to a server and receiving the result is intolerable

\subsection{Other workflows}
Although most of the analysts we spoke to use and appreciates the sharing
features, RCloud is less popular as a tool for exploratory data analysis and
development.

Much of this is because of familiarity with other tools. Horace: ``for other
people who use RCloud vs RStudio, they will always think that they want to
prototype on RStudio. Maybe because it's more familiar for them. So, they work
on RStudio, on a small data set, and they fix up everything, make sure
everything is right, and then they place it on RCloud for sharing.''

Rick says, ``I find that the style of editing is clunky. I would never edit like that
if I'm doing ordinary work. I usually use vi.'' ``I mean it's nice to have it
saved, but there's this trade-off between it making it easier for me to present
something or to save something, and my ease of typing and correcting and things
in a plain old editor window.''

Many of the analysts we spoke to preferred using a text editor to refine and
store their work, pasting it to an R command line to try it out. Rick describes,
``I have the discipline of having a file that describes what I'm doing with the
commands that I'm using so that I can go back and recreate it or pass it on to
someone else, so it's a little bit like having an RCloud notebook, it's not
necessarily executable but most of the commands that I've typed are in there.''

Kenny works much the same way, rearranging the file so the good code is at the top of
the file. Then he saves the file once he's got the right code at
the top of the file (and often shares the work by pasting it into an RCloud notebook).

Taylor asks, ``why would I want to use RCloud over my
current setup? If it's just me, I like my text editor and terminal. There's
nothing that I want that those two don't give me.''

Working in a shared environment also entails compromises about what you can
install. Byers says that installing random packages is part of exploratory data
analysis.

\subsection{Proliferation of notebook cells}
RCloud's notebook interface combines editable Markdown with a command line
interface.

The interface takes some learning. Clint says ``it took me a couple of weeks of
looking at it to become comfortable because I saw the cells and it just threw me
for a total loop. I mean, it's a good idea because [...] I can run it all at
once if I want or I can break these into sections for either debugging or
staging purposes. I really like it now, but when I first saw it, [it was] very
confusing [...] I think I prefer it now, because you can cut your code up.''

Just the difference between a cell and the command prompt can confuse and slow
down users. Rick: ``if I was typing into one of the cells near that top, I had
to think `I'm editing this cell and then I have to execute it', and if was doing
the one at the bottom I could type it but then it would automatically execute
[by pressing enter], but then it became a standard cell and I have to edit
it. It was just another set of modes that I didn't like''

For some users, the cells getting saved feels incompatible with exploratory data
analysis. Rick notes that Markdown is a form of literate programming, and
``being literate about anything typically takes a lot of rewriting and going
back over things [...] I just want to explore a few things and then I'll know
what I want to write.''

Much of the time, commands typed at the R command line just serve a transitory
purpose, so having RCloud save commands into cells bothers experienced
users. Kenny notes, ``It saves everything I do like everything is gold, but most
of it is junk not meant to be saved.'' Rick says ``I just want to type in a
couple of quick and get some results that are going to tell me what to do next,
and they're not necessarily archival in any sense''

The material that is not appropriate to save includes ``expressions that allow
me to check that I'm the right track'' (Rick), ``checking out what your data is,
or you make a plot of the data. Things that should not really become part of a
notebook, but things that help you understand your data better.'' (Raif)

Users also felt that cells do not capture the right level of granularity: they
either hold too much code or too little. Rick says that RCloud's cells structure
``tempts me to type a big long thing and then run the whole thing, as opposed to
typing a few little pieces that I'm thinking of on the inner parts and then put
them together to make the big thing'' as he would do on the command line. When
Taylor uses the R command line, he copies and pastes ``like 5-10 lines of code,
so when something breaks, I get an error message on that one line, and I can
up-arrow and change it and fix it, whereas in RCloud I have to run a whole cell,
so the only way to get that same functionality is if every line's in one
cell. Because if it runs lines 1, 2, 3, then 4 fails, I have to either run the
whole cell again, or have line 4 in its own cell. I can't rerun just one line.''

Cells can also take up substantial vertical space, requiring scrolling. Rick
complains that ``every time I type more stuff, the notebook gets longer and
longer and it's harder to deal with.'' Ivan says that cells' controls and blank
space take extra real estate compared to a straight command line, and he and
KCKC both complained that when there are long results or plots, it causes the
code to scroll off the screen. Raif: ``for example imagine I run a cell whose
output is huge, and now that output would block the whole window, and in order
to find the next cell, I have to scroll down, and find where that cell starts,
right? So, I'm losing the continuity of my code to the output''

Some users think users should have the option to keep results separate from the
code.  Ivan thinks that RStudio's layout is more helpful because the charts
are kept in another pane and stay in one place while doing analysis. Raif says
``cells are really useful'' but ``you want to see your output in a different
window [or] on a separate part of the window''.

The cell structure also can be problematic if some cells take a long time to
run. In Kenny's work, there is often a ``long tail''.  The first cell may take
seconds to run, the next cell minutes, and the last cell 5 hours.  In this
situation, the ``Run whole notebook'' button is ``dangerous''. When converting
his work to notebooks, Kenny ends up with a lot of comments that say ``this cell
takes a long time to run.'' This is also a huge
issue for demoing code.

The solution is probably something like knitr's caching of results, which Kenny
is doing in an ad-hoc way, ``littering the notebook with little switches that
comment out'' the slow parts and ``load the object instead'', ``save the object
to disk''. One of the authors of this paper manually writes cells that check if
a result file exists, and perform the calculation only if it's not there.

@@

RCloud combines the functions of scratchpad with notebook assets. KC thinks
that having a scratch pad that doesn't persist between different notebooks is
``useless''.

%% The one-window design also makes some users feel that the side panes are taking
%% up too much space. Ivan: having ``long term'' code (which would be in assets)
%% above the command prompt, like in RStudio, means that both can be sized very
%% wide to accomodate long lines.

%% Kenny: Problem with screen real estate: side panes take up too much
%% space. Doesn't need the notebook tree pane.

%% Corey: In the previous version, it switched between code and results. Now you
%% see both - ``and there are times when that's good, but there's certainly times
%% when it'd be nice to - you know, you have 10 or 15 cells and you just want to
%% get back and there's all this output and you're scrolling all over the place so,
%% if there were a way to control both, that'd be great. There are times when I do
%% want to see the output, but there are times when I've seen enough of the same
%% thing.''

%% Jyothi: ``It's good that I can see the results but if I'm working and I need
%% more room and I want to hide the results, I can't.'' Could we have ``an option
%% to close the results window?''

\subsection{A sea of notebooks}
RCloud is starting to be a victim of its own success, as there are now too many
notebooks.  Rick doesn't find the notebook tree very useful for this reason: ``I
don't necessarily need to see everybody's notebook that uses RCloud'' ``every
time I do something new, I get a new notebook and so now my notebooks are maybe
50 or 60. That's enough to think about just on my own, but if I've got
everybody's 50 or 60 sitting on my display, I find that it's more than I want to
know about.''

Although RCloud promises an environment where notebooks stay working, our users
have not all learned the habits that make this a reality. As Byers puts it,
there is still a big problem with ``bitrot'' - notebooks often stop working
because the user changed the structure of their data, changed a filename or a
database.  He says we need ``organizational protocols'' to catch up with the
technology.

There are technical problems to be solved here as well. Byers continues that the
problem with forking a notebook and correcting it, is that the original notebook
still exists with the error.  We need a way to see that there is a more recent
fork of notebook and to fetch in the changes from it.

Taylor, who writes tools and example notebooks for those tools, seems to be
haunted the most by dead notebooks.  At first, he had to learn how to write safe
notebooks: ``It would be a file inside a directory that only I had [...] access
to and so someone else couldn't run it, [...] a temporary file on something that
was read-only to everyone else. But I've gotten careful about that.''

But even correcting notebooks can come too late: ``I'll make a notebook of a
little tutorial of basic functions and what they do, and then I'll send [it]
out, and then people will give me feedback on [it], or have some new feature, or
I'll realize something's wrong, and then I'll go back and change it and update
that page, but then what happens is people have in the meantime forked it. [...]
Usually what it'll be is actually that I'll have to change a package as well, so
their old fork stops working, and then they complain. [...] This probably bites
me more than other people, because I'm doing development so things are changing
as they're looking at the code. Like, iotools, a lot of the functions change
names or change syntax.''

Taylor tried to keep his notebook tree tidy, but this didn't help either, because
notebooks still exist when they have left the tree. ``That's just a problem because people will have old notebooks that
don't work anymore. It's annoying when it's a notebook that I still have and the
format has changed just a bit. More than that is when people have, now that I
know that you can't even delete notebooks, they'll have a fork of something that
doesn't exist anymore on my thing.'' ``Now I'm really gun-shy about sharing
notebooks because it's like, do I want to support this forever?''

Taylor: ``A good example is, someone was complaining about doing forking in an
RCloud notebook. They were trying to use a parallel package. And I said, no no I
think it works, and I did like a one-cell thing that just did some sweep kind of
thing to show that it was actually working in parallel, and then I just deleted
it, thinking it was gone. And then it saved something in a directory that I then
deleted, because it was just a temporary directory. Three months later, I don't
know why they were running it, they said `oh it doesn't work'. I think that's
the biggest issue I'm having right now, all these forks.''

Taylor and Horace work together on a notebook. ``There's no way for both of us
to have ownership of a notebook, so the only way is to fork it back and forth,
and so we have dozens of old copies. [We end up] deleting all the old ones, but
people still have links to them, because they don't actually disappear. [...]
The UI makes it easy to go back to the old stale versions, because you can say
`forked from' `forked from' `forked from', but there's actually no way to go the
way that we want people to go, which is, like, up the tree.''

%% Bart: ``One UI notebook and a bunch of other notebooks that do various tasks and
%% are called with API calls. The problem is, if you clone all the notebooks, then
%% you have to go update every single notebook, because all the IDs are different
%% now. There's no sort of relative path type stuff that you can do. You can do it
%% with static assets, but you can't do it with Ajax calls''
